Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Updated Basics.

In this paper, a {\em geometric constraint system} is a multivariate 
polynomial (usually bilinear or quadratic) system
$G(x,\delta)=0$, representing constraints with parameters $\delta$ 
between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
%
When the type of constraint (system) is fixed, 
the system is simply represented as $(G,\delta)$, where 
$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
with the vertices $V$ representing the geometric primitives in 
$\mathbb{R}^2$ and (hyper)edges
$E$ representing the constraints, each with an associated 
parameter $\delta$.

For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
$G=(V,E)$  
with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
this represents the distance constraint system $\| x_u -x_v \|_2 =
\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
$u\in V$.

In this paper, we will consider 3 types of constraint systems: bar-joint
as well as {\em pinned line-incidence}, and {\em body-hyperpin}. 
In all of these cases, a cartesian \emph{realization} or 
\emph{solution} $G(p)$ of $(G,\delta)$ is 
an assignment of coordinates or Euclidean transformations (poses), 
$p: V \rightarrow
\mathbb{R}^2$ or $\mathbb{R}^3$ to the vertices of $G$ satisfying the 
constraints with parameters $\delta$, 
modulo orientation preserving isometries 
(Euclidean rigid body motions except for pinned line-incidence).

Although the realization space itself depends on the constraint parameters 
$\delta$, many  relevant {\em generic} properties of 
the constraint system $G(x,\delta)$
are properties of the constraint (hyper)graph $G$ 
and do not depend on $\delta$ (or they hold for all but a measure 
zero set of $\delta$
values). Many of these are properties of the jacobian 
$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of
$G$} 
(a matrix of indeterminates).
For example, the {\em bar-joint rigidity matrix 
of the graph $G = (V,E)$}
is a matrix of indeterminates representing the 
jacobian of the distance map 
$\| x_u -x_v \|_2$  for $(u,v) \in E$.
The matrix has one row per edge $(u,v)$ in $E$,  
with 2 columns for $u$ (resp.
$v$) containing the 2 coordinate indeterminates 
for $x_u -x_v$ (resp. $x_v-x_u$), 
that is 4 non-zero entries per row.

Some of these generic constraint system or 
(hyper)graph properties are the following ({\bf Note:} we refer to these as
properties of the constraint system or as properties of the underlying
(hyper)graph interchangeably).
%
A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
has independent rows
(i.e, the determinant of some square submatrix is not
identically zero). 
It is {\em isostatic (minimally rigid, well-constrained)} 
if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.

@@

@@

\todo{Where to put these}
A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.
The left nullspace of the rigidity matrix is called the {\it space of stress}. A vector in the stress space is called a {\it flex vector}.

>>>>

In this paper, a {\em geometric constraint system} is a multivariate polynomial 
(usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ 
between geometric primitives  in $\mathbb{R}^2$ represented collectively $x\in \mathbb{R}^n$.   
%
When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where 
$G$ is the underlying constraint (hyper)graph $G = (V,E)$ 
with the vertices $V$ representing the geometric primitives in 
$\mathbb{R}^2$ and (hyper)edges
$E$ representing the constraints, each with an associated 
parameter $\delta$.
%
For example, a {\em bar-joint system or linkage} $(G,\delta)$, is a graph
$G=(V,E)$  
with fixed length bars as edges, i.e. $\delta: E \rightarrow \mathbb{R}$;
this represents the distance constraint system $\| x_u -x_v \|_2 =
\delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of
$u\in V$.

In this paper, we will consider 3 types of constraint systems: bar-joint
as well as {\em pinned line-incidence}, and {\em body-hyperpin}. 
In all of these cases, a cartesian \emph{realization} or 
\emph{solution} $G(p)$ of $(G,\delta)$ is 
an assignment of coordinates or Euclidean transformations (poses), 
$p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the 
constraints with parameters $\delta$, 
modulo orientation preserving isometries 
(Euclidean rigid body motions for distance constraint).

Although the realization space itself depends on the constraint parameters 
$\delta$, many  relevant {\em generic} properties of 
the constraint system $G(x,\delta)$
are properties of the constraint (hyper)graph $G$ 
and do not depend on $\delta$ (or they hold for all but a measure 
zero set of $\delta$
values). Many of these are properties of the jacobian 
$\Delta_x G(x,\delta)$, often called the appropriate {\em rigidity matrix of
$G$} 
(a matrix of indeterminates).
For example, the {\em bar-joint rigidity matrix 
of the graph $G = (V,E)$}
is a matrix of indeterminates representing the 
jacobian of the distance map 
$\| x_u -x_v \|_2$  for $(u,v) \in E$.
The matrix has $2$ columns per vertex in $V$ and one row per edge  in $E$, 
where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates 
for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), 
that is 4 non-zero entries per row.

Some of these generic constraint system or 
(hyper)graph properties are the following ({\bf Note:} we refer to these as
properties of the constraint system or as properties of the underlying
(hyper)graph interchangeably).
%
A constraint (hyper)graph $G$ is {\em independent} if its appropriate rigidity matrix of indeterminates  
has independent rows
(i.e, the determinant of some square submatrix is not
identically zero). 
It is {\em rigid} if the number of independent rows of the rigidity matrix is
at least the number of columns less the number of rigid body motions, 
which is 3 for distance constraint systems. 
It is {\em isostatic (minimally rigid, well-constrained)} 
if it is both rigid and independent. 
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$, 
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained. 
It is {\em flexible} if it is not rigid, {\em under-constrained} if it is independent and not rigid, or {\em over-constrained} if it is not independent.

@@

%Define rigidity matroid and combinatorial characterization of independence/rigidity next
%Laman's theorem (for bar-joint rigidity matroid), 
%cite also Lovasz Yemini
%Cite other combinatorial characterizations for other types of constraint systems/hyper graphs:
%Tay and Whiteey, White and Whitelely,  
%Streinu Theran (k,l)-sparsity matroid

The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$, 
where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid, 
if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.

There exist various results on combinatorial characterization of independence and rigidity
for different types of  rigidity matroid. 
For bar-joint rigidity matroid, 
the famous Laman's theorem~\cite{???} states that the underlying graph
is minimally rigid if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. 
The result by Lovasz and Yemini~\cite{???} shows that all 6-vertex-connected graphs are rigid in the plane. 
For bar-body rigidity matroid,
Tay~\cite{???} proved that
the underlying multigraph is minimally rigid if and only if it can be decomposed as $3$ edge disjoint spanning trees. 
White and Whiteley~\cite{???} gave the same characterization using a different technique to study the algebraic - geomtric conditions called pure condition.
More generally, 
Lee, Streinu and Theran~\cite{???} defined the $(k,l)$-sparsity matroid,
where a hypergraph is called $(k,l)$-sparse if $|E'| \le k|V| - l$ for any induced subgraph with at lease 2 vertices.

@@

For bar-joint graphs,  
a vector  in the right null space of its rigidity matrix
is called a {\em flex vector}. 
It has 2 entries of indeterminates per vertex and represents the internal motions of the system
(i.e.\ modulo rigid body motions).
For rigid graphs, all flex vectors are  identically zero. 
A vector in the left null space of the rigidity matrix
is called a {\em stress vector}, with one entry of indeterminate per edge.
For independent graphs, all stress vectors are identically zero.
